---
title: "SOX2 Copy Number Variation and TP53 Mutation: Fisher’s Exact Test Analysis"
author: "Siber Gunes Celik"
output: html_document
---

## Abstract

This project investigates the association between SOX2 copy number variation and TP53 mutation status across various cancer types using Fisher’s Exact Test. Data was obtained from [cBioPortal](https://www.cbioportal.org), and results are visualized and interpreted using R and relevant packages.

## Data Source

Data was downloaded from [cBioPortal](https://www.cbioportal.org), including patient IDs, SOX2 copy number variation, and TP53 mutation status.

## Data Preparation

The columns were cleaned and renamed for easier manipulation, and a contingency table was created to summarize the co-occurrence of the mutations.

``` {r, message=FALSE, warning=FALSE}
library(ggstatsplot)
library(RColorBrewer)
```

``` {r}
# Load and prepare data
my_data <- read.delim('pancancer_combined.txt')
colnames(my_data) <- c('ID', 'TP53_status', 'SOX2_status')
head(my_data)
```

## Statistical Test

We used Fisher’s Exact Test to assess the association between SOX2 copy number gain and TP53 mutation status. Visualizations include mosaic plots and bar charts.

``` {r}
# Define the attributes of interest 
atr1 <- "Missense"
atr2 <- "Gain"

# Subsetting data to generate contingency table
Gene1atr1 <- subset(my_data, subset= TP53_status == atr1)
Gene2atr2 <- subset(my_data, subset = SOX2_status == atr2)

Cont1 <- subset(Gene1atr1, subset = SOX2_status == atr2)
Cont2 <- subset(Gene1atr1, subset = SOX2_status != atr2)
Cont3 <- subset(Gene2atr2, subset = TP53_status != atr1)
Cont4 <- subset(my_data, subset = TP53_status != atr1)
Cont4 <- subset(Cont4, subset = SOX2_status != atr2)

# Create contingency table
dat_table_2 <- data.frame(
  "atr1_present" = c(length(Cont1$ID), length(Cont3$ID)),
  "atr1_absence" = c(length(Cont2$ID), length(Cont4$ID)),
  row.names = c("atr2_present", "atr2_absence"),
  stringsAsFactors = FALSE
)
colnames(dat_table_2) <- c("atr1_present", "atr1_absence")
dat_table_2

# Contingency Table Visualization
mosaicplot(dat_table_2,
           main = "Mosaic plot of Contingency Table",
           color = c('#FFE4E1', '#BC8F8F'),
           border = TRUE
)

# Fisher`s Exact Test

test_2 <- fisher.test(dat_table_2)
test_2

```

## Results

Fisher's Exact Test yielded a p-value < 0.001 and an odds ratio of 2.47, suggesting a statistically significant association between SOX2 gain and TP53 mutation in the dataset.


``` {r}
# Create a transposed data frame for ggbarstats plot
x_2 <- c()
for (row in rownames(dat_table_2)) {
  for (col in colnames(dat_table_2)) {
    x_2 <- rbind(x_2, matrix(rep(c(row, col), dat_table_2[row, col]), ncol = 2, byrow = TRUE))
  }
}
df_2 <- as.data.frame(x_2)
colnames(df_2) <- c("gene1_atr1", "gene2_atr2")
head(df_2)

# Plotting results with ggbarstats
test_2 <- fisher.test(table(df_2))
ggbarstats(
  df_2, gene1_atr1, gene2_atr2,
  results.subtitle = FALSE,
  package = "RColorBrewer",
  palette = "Paired",
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test_2$p.value < 0.001, "< 0.001", round(test_2$p.value, 3))
  )
)

```


## References

1: Oliveira, Natalia L., et al. “A discussion on significance indices for contingency tables under small sample sizes.” PLoS One 13.8 (2018): e0199102.

2: Kim, Hae-Young. “Statistical notes for clinical researchers: Chi-squared test and Fisher’s exact test.” Restorative Dentistry & Endodontics 42.2 (2017): 152-155.




